
\begin{table}[H]
\begin{tabular}{|c|ccccc|}
\hline
\textbf{Color space} & \textbf{AP Airplanes} & \textbf{AP Cars} & \textbf{AP Faces} & \textbf{AP Motorbikes} & \textbf{MAP}\\
\hline
gray & 0.8564 & 0.7468 & 0.6014 & 0.8485 & 0.7633\\
normalizedRGB & 0.9245 & 0.8081 & 0.8906 & 0.8308 & 0.8635 \\
RGB & 0.8870 & 0.8407 & 0.6821 & 0.8815 & 0.8228 \\
opponent & 0.8918 & 0.7946 & 0.8918 & 0.8457 & 0.8560\\
\hline
\end{tabular}
\caption{Effect of different color spaces for key SIFT}
\label{tab:key}
\end{table}

As can be seen in table \ref{tab:key}, the normalized RGB color space shows the best results for automated image classification in comparison with the other color spaces when key-point SIFT is used. This result can be contributed to the fact that this color space is scale-invariant and invariant to light intensity changes. The opponent color space is also shift-invariant with respect to light intensity and shows a similar, but slightly lower, performance as the normalized RGB color space. The gray and RGB color spaces do not exhibit invariant features and therefore perform worse than the Opponent and normalized

and outperforms the classifiers which use gray or RGB color space, which are not invariant.
\cite{van2010evaluating}

\begin{table}[H]
\begin{tabular}{|c|ccccc|}
\hline
\textbf{Color space} & \textbf{AP Airplanes} & \textbf{AP Cars} & \textbf{AP Faces} & \textbf{AP Motorbikes} & \textbf{MAP}\\
\hline
gray & 0.6476 & 0.7741 & 0.4292 & 0.8165 & 0.6668\\
normalizedRGB & 0.5743 & 0.3402 & 0.4678 & 0.5646 & 0.4867\\
RGB & 0.6424 & 0.9726 &  0.8689 & 0.6747 & 0.7897\\
opponent & 0.6447 & 0.9053 & 0.9510 & 0.7516 & 0.8132\\
\hline
\end{tabular}
\caption{Effect of different color spaces for dense SIFT}
\label{tab:color_sift}
\end{table}


Table \ref{tab:color_sift} shows that Opponent SIFT has the highest Mean Average Precision score. The opponent color space is shift-invariant with respect to light intensity and outperforms the classifiers which use gray or RGB color space, which are not invariant. Surprisingly, the classifier which uses the normalized RGB color space performs worst, even though this color space is scale-invariant and invariant to light intensity changes. As the images of the dataset are mostly captured from the same point of view and have similar scales, the scale-invariance feature of the normalized RGB does not significantly contribute to the automated image classification. The normalization of RGB possibly discards more distinctive features than the other color space SIFT methods and therefore performs worse.\\